### 2015-11-19 CV/Brian Goldman CV.tex
Fixed spacing
Minor space saving
Split 'peer reviewed' from 'other' publications
Some forgotten edits to my application materials

\documentclass[a4paper,11pt]{article}
\usepackage{cv}
\usepackage{verbatim}
\name{Brian W. Goldman}
\info{Address: & Department of Computer Science \& Engineering, Michigan State University, East Lansing, Michigan.\\
      Phone: & 314-313-1281\\
      Email: & brianwgoldman@acm.org\\
      Website: & \url{http://brianwgoldman.github.io/}}

@@

\section{Education}
\begin{itemize}
\item 2015 Doctor of Philosophy in Computer Science \& Engineering, Michigan State University
\begin{itemize}
  \item Dissertation Advisor:  Dr. William F. Punch%, GPA: 4.0/4.0
  \item Dissertation: Out of the Box Optimization using the Parameter-less Population Pyramid
\end{itemize}
\item 2012 Master of Science in Computer Science, Missouri University of Science \& Technology
\begin{itemize}
  \item Thesis Advisor:  Dr. Daniel R. Tauritz%, GPA: 4.0/4.0
  \item Thesis: Robust Evolutionary Algorithms
\end{itemize}
\item 2010 Bachelor of Science in Computer Science, Missouri University of Science \& Technology
%\begin{itemize}
%  \item Minor in Mathematics, GPA: 3.84/4.0 in major, 3.81/4.0 overall
%\end{itemize}
\end{itemize}
\section{Academic Positions}
\begin{itemize}
\item Research Assistant Professor, Colorado State University. Starting Spring 2016.
\item Postdoc, Hintze Lab, Michigan State University. Fall 2015.
\item Advanced Computing Solutions Program, Los Alamos National Laboratories. Summer 2012.
\item Center for Cyber Defenders, Sandia National Laboratories. Summer 2011.
\end{itemize}

\begin{publications}
%\printbib{books}
\printbib{journals}
\printbib{papers}
\printbib{posters}
\printbib{curpapers}
%\printbib{chapters}
%\printbib{techreports}
%\printbib{bookreviews}
%\printbib{editorials}
\end{publications}

\section{Grants}
\begin{itemize}
\item BEACON NSF STC 2015 funded proposal \$105,148
\item BEACON NSF STC 2013 funded proposal \$62,284.
\end{itemize}

@@

\section{Invited talks}
\begin{itemize}
\item Departmental Seminar, University of Memphis, April 2015
\item BEACON NSF STC Congress, Michigan State University, August 2014 and August 2015
\item Evolutionary Computation, Michigan State University, September 2014 and October 2015
\item Heuristic and Evolutionary Algorithms Laboratory, University of Applied Sciences Upper Austria, December 2014
\item BEACON Day, North Carolina A\&T, October 2013
\end{itemize}

@@

\section{Teaching Experience}
\begin{itemize}
\item Introduction to Programming II, Summer 2014.
      Lead instructor for C++11 course. In charge of lecturing, creating course content, and assessments
      for 37 students. Managed two graduate assistants who taught lab sections and graded assignments.
      Received 3.81/4 student evaluation.
\item Introduction to Programming II, Fall 2013 -- Spring 2014. Primary instructor for lab section.
\item Object Oriented Numerical Methods, Spring 2012. Teaching assistant for upper level intensive C++ course.
\item Introduction to C++ Lab, Spring 2011 -- Fall 2011. Primary instructor for lab section.
\item Parallel Computing, Spring 2011. Teaching assistant for upper level course.
\item Evolutionary Computation, Fall 2010. Teaching assistant for upper level course.
\end{itemize}

>>>>

\documentclass[a4paper,11pt]{article}
\usepackage{cv}
\usepackage{verbatim}
\setlist[itemize,enumerate]{noitemsep,nolistsep}
\name{Brian W. Goldman}
\info{Address: & Department of Computer Science \& Engineering, Michigan State University, East Lansing, Michigan.\\
      Phone: & 314-313-1281\\
      Email: & brianwgoldman@acm.org\\
      Website: & \url{http://brianwgoldman.github.io/}}

@@

\section{Education}
\begin{itemize}
\item 2015 Doctor of Philosophy in Computer Science \& Engineering, Michigan State University
\begin{itemize}
  \item Dissertation Advisor:  Dr. William F. Punch%, GPA: 4.0/4.0
  \item Dissertation: Out of the Box Optimization using the Parameter-less Population Pyramid
\end{itemize}
\item 2012 Master of Science in Computer Science, Missouri University of Science and Technology
\begin{itemize}
  \item Thesis Advisor:  Dr. Daniel R. Tauritz%, GPA: 4.0/4.0
  \item Thesis: Robust Evolutionary Algorithms
\end{itemize}
\item 2010 Bachelor of Science in Computer Science, Missouri University of Science and Technology
%\begin{itemize}
%  \item Minor in Mathematics, GPA: 3.84/4.0 in major, 3.81/4.0 overall
%\end{itemize}
\end{itemize}
\section{Academic Positions}
\begin{itemize}
\item Research Assistant Professor, Colorado State University. Starting Spring 2016.
\item Postdoc, Hintze Lab, Michigan State University. Fall 2015.
\item Advanced Computing Solutions Program, Los Alamos National Laboratories. Summer 2012.
\item Center for Cyber Defenders, Sandia National Laboratories. Summer 2011.
\end{itemize}

\begin{peer-publications}
%\printbib{books}
\printbib{journals}
\printbib{papers}
%\printbib{chapters}
%\printbib{techreports}
%\printbib{bookreviews}
%\printbib{editorials}
\end{peer-publications}

\begin{other-publications}
\printbib{posters}
\printbib{curpapers}
\end{other-publications}

\section{Grants}
\begin{itemize}
\item BEACON NSF STC 2015 funded proposal \$105,148
\item BEACON NSF STC 2014 funded proposal \$62,284.
\end{itemize}

@@

\section{Invited talks}
\begin{itemize}
\item Job Talk, University of Memphis, April 2015
\item BEACON NSF STC Congress, Michigan State University, August 2014 and August 2015
\item Evolutionary Computation, Michigan State University, September 2014 and October 2015
\item Heuristic and Evolutionary Algorithms Laboratory, University of Applied Sciences Upper Austria, December 2014
\item BEACON Day, North Carolina A\&T, October 2013
\end{itemize}

@@

\section{Teaching Experience}
\begin{itemize}
\item Introduction to Programming II, Summer 2014.
      Lead instructor for a C++11 course. In charge of lecturing, creating course content, and assessments
      for 37 students. Managed two graduate assistants who taught lab sections and graded assignments.
      Received 3.81/4 student evaluation.
\item Introduction to Programming II, Fall 2013 -- Spring 2014. Primary instructor for lab section.
\item Object Oriented Numerical Methods, Spring 2012. Teaching assistant for senior level course.
\item Introduction to C++ Lab, Spring 2011 -- Fall 2011. Primary instructor for lab section.
\item Parallel Computing, Spring 2011. Teaching assistant for senior level course.
\item Evolutionary Computation, Fall 2010. Teaching assistant for senior level course.
\end{itemize}

### 2015-11-19 ResearchStatement/main.tex
Fixes from Charles' suggestions
Some forgotten edits to my application materials

\noindent
For many problems in many disciplines, it is easier to determine the quality of a solution
than to find the best possible solution. For example, this is true of NP-Hard problems,
finite element analysis, chemical design, and physics simulations.
While coming from very different fields, these problems share
three critical characteristics:
\begin{itemize}
\item There is a defined set of possible solutions.
\item We can measure the quality of a proposed solution.
\item There is no tractable method for finding the best solution.
\end{itemize}
I design algorithms for optimizing problems with these characteristics. Instead of
attempting to create specialized heuristics for every problem,
my goal is to discover and utilize general principles which can
then be applied to a large set of problems without further expert knowledge.

@@

The primary focus of my dissertation was the creation of the
Parameter-less Population Pyramid (P3), an optimization algorithm
that requires no problem specific configuration~\cite{goldman:2014:p3,goldman:2015:fastp3}.
P3 is comprised of three parts: Stochastic local search, entropy based model building, and
population layering. The guiding principle behind its design is to exploit all previous
knowledge the algorithm has about the problem before resorting to further stochastic exploration.
Each iteration, P3 creates a new high quality solution using local search and a model of
problem structure. The model is built by examining the mutual information between variables
in previously found high quality solutions~\cite{goldman:2012:ltga}.
By storing multiple populations of solutions, P3 is able to progressively guide search
to better and better areas without becoming stuck or requiring large initial random search.
P3 is able to find the global optimum in less time than other techniques, as well as
maintain a better ``current best'' solution as optimization progresses across a diverse set of problems.
This quality is on top
of the fact that the comparison algorithms used unrealistically ideal configurations while
P3 requires no configuration.

@@

Determining the placement of turbines in a wind farm is another candidate for
generic optimization. Aerospace engineers have developed models which can calculate
how the wake of each turbine can effect the electrical output of the entire field.
While turbines are commonly arranged in grids, we found using optimization methods
that fields with less regular layouts can be significantly more efficient.

A challenging task in machine learning is the training of large scale neural networks.
I have recently been in collaboration with an evolutionary biologist
who studies natural learning to develop better methods for training neural networks.

\section{Future Work}
In recently submitted work~\cite{whitley:2016:mkl}, we determined that many real-world
instances of common NP-Hard problems contained a surprising amount of regularity.
I believe this regularity can be exploited to significantly reduce the space
which needs to be searched for high quality solutions. To that end, I plan to develop
a problem independent technique for search space compression.

A current limitation of P3 is the focus on single objective optimization. For many
problems, there are many competing objectives which should be optimized. For example,
strength, weight, and cost in an engineering task. Optimization gains significant
benefits from treating this problem as having three separate goals instead of attempting
to combine them into a single objective. I believe that integrating P3 with the concepts
of multi-objective optimization will result in a powerful tool for real-world optimization.

>>>>

\noindent
For many problems in many disciplines, it is easier to determine the quality of a solution
than to find the best possible solution. For example, this is true of most NP-Hard problems,
finite element analysis, chemical design, and physics simulations.
While coming from very different fields, these problems share
three critical characteristics:
\begin{itemize}
\item There is a defined set of possible solutions.
\item We can measure the quality of a proposed solution.
\item There is no tractable method for finding the best solution.
\end{itemize}
I design algorithms for optimizing problems with these characteristics. Instead of
attempting to create specialized heuristics for every problem,
my goal is to discover and utilize general principles which can
then be applied to a large set of problems without further expert knowledge.

@@

The primary focus of my dissertation was the creation of the
Parameter-less Population Pyramid (P3), an optimization algorithm
that requires no problem specific configuration~\cite{goldman:2014:p3,goldman:2015:fastp3}.
P3 is comprised of three parts: Stochastic local search, entropy based model building, and
population layering. The guiding principle behind its design is to exploit all previous
knowledge the algorithm has about the problem before resorting to further stochastic exploration.
Each iteration, P3 creates a new high quality solution using local search and a model of
problem structure. The model is built by examining the mutual information between variables
in previously found high quality solutions~\cite{goldman:2012:ltga}.
By storing multiple populations of solutions, P3 is able to progressively guide search
to better and better areas without becoming stuck or requiring large initial random search.
P3 is able to find the global optimum in less time than other techniques, as well as
maintain a better ``current best'' solution as optimization progresses, across a diverse set of problems.
This quality is on top
of the fact that the comparison algorithms used unrealistically ideal configurations while
P3 requires no configuration.

@@

Determining the placement of turbines in a wind farm is another candidate for
generic optimization. Aerospace engineers have developed models which can calculate
how the wake of each turbine effects the electrical output of the entire field.
While turbines are commonly arranged in grids, we found using optimization methods
that fields with less regular layouts can be substantially more efficient.

A challenging task in machine learning is the training of large scale neural networks.
I have recently been in collaboration with an evolutionary biologist
who studies natural learning to develop better optimization methods for training neural networks.

\section{Future Work}
In recently submitted work~\cite{whitley:2016:mkl}, we determined that many real-world
instances of common NP-Hard problems contained a surprising amount of regularity.
I believe this regularity can be exploited to reduce the space
which needs to be searched for high quality solutions. To that end, I plan to develop
a problem independent technique for search space compression.

A current limitation of P3 is the focus on single objective optimization. For many
problems, there are many competing objectives which should be optimized. For example,
strength, weight, and cost in an engineering task. Optimization gains substantial
benefits from treating this problem as having three separate goals instead of attempting
to combine them into a single objective. I believe that integrating P3 with the concepts
of multi-objective optimization will result in a powerful tool for real-world optimization.

### 2015-11-19 TeachingStatement/main.tex
Some forgotten edits to my application materials

\section{Teaching Classes of Moderate Size}
Due to my previous success as a Lab Instructor, I was offered a
position as Lead Instructor for the second level undergraduate
programming course. This class is designed to introduce the students,
who learn Python in the first semester, to C++11. They learn
everything from basic syntax and STL tools to templates and how
to building their own data structures with dynamic memory.
I was given full autonomy to create course content, including
lectures, assignments, and assessments. To help with grading
assignments and lab work created by the 37 students, the
department assignment me two Lab Instructors.
I am proud to report that I received an average
of 3.81 out of 4 in post course student assessment.

In order to create an inclusive environment, I worked to
show my concern for each student's understanding. While
lecturing I frequently paused and prompted for questions,
and did my best to gauge whether students were understanding
the material. This would often result in me changing the pace
of the lecture to meet the students' comprehension, either
skipping slides or moving to the chalkboard to give further
examples. For those students who did not want to ask questions
in class, I utilized Piazza, an online tool where students
can post their questions anonymously. These questions could be
visible to everyone in the class, allowing students to answer
each others' questions, or only visible only to me.

@@

\section{Teaching Classes of Large Size}
While my primary experiences have been with moderately sized
classes (25-50 students), I have taught a few lectures to much
larger classes (120+). Lectures of that size make it much more
difficult to interactively gauge student comprehension, making
alternative teaching tools all the more important. The small
group worksheet approach became even more critical as, unlike
the lecture, it is unaffected by class size. It also provided
me an opportunity to work with small groups of students to
answer specific questions or to clarify ideas, something impossible
during large scale lecture. As a teaching assistant in these
courses I saw the effectiveness of video tutorials and online
classroom management systems which allow students to work together to answer each other's questions.

@@

\section{Future Goals}
While much of my teaching experience has been with introductory
programming courses, I am also interested in teaching higher level
and theory based courses. Artificial intelligence and machine learning
are common courses which coincide strongly with my research interests.
Algorithms, discrete math, and graph theory are topics that I believe
present exciting challgenges for keeping students engaged.
From previous work I have also acquired significant knowledge of
introductory and high performance C++ and would enjoy teaching classes
on those topics.

>>>>

\section{Teaching Medium-Sized Classes}
Due to my previous success as a Lab Instructor, I was offered a
position as Lead Instructor for the second level undergraduate
programming course. This class is designed to introduce the students,
who learn Python in the first semester, to C++11. They learn
everything from basic syntax and STL tools to templates and how
to building their own data structures with dynamic memory.
I was given full autonomy to create course content, including
lectures, assignments, and assessments. To help with grading
assignments and lab work created by the 37 students, the
department assignment me two Lab Instructors.
I am proud to report that I received an average
of 3.81 out of 4 in post course student assessment.

In order to create an inclusive environment, I worked to
show my concern for each student's understanding. While
lecturing I frequently paused and prompted for questions,
and did my best to gauge whether students were understanding
the material. This would often result in me changing the pace
of the lecture to meet the students' comprehension, either
skipping slides or moving to the chalkboard to give further
examples. For those students who did not want to ask questions
in class, I utilized Piazza, an online tool where students
can post their questions anonymously. These questions could be
visible to everyone in the class, allowing students to answer
each other's questions, or only visible only to me.

@@

\section{Teaching Large Classes}
While my primary experiences have been with moderately sized
classes (25-50 students), I have taught a few lectures to much
larger classes (120+). Lectures of that size make it much more
difficult to interactively gauge student comprehension, making
alternative teaching tools all the more important. The small
group worksheet approach became even more critical as, unlike
the lecture, it is unaffected by class size. It also provided
me an opportunity to work with small groups of students to
answer specific questions or to clarify ideas, something impossible
during large scale lecture. As a teaching assistant in these
courses I saw the effectiveness of video tutorials and online
classroom management systems which allow students to work together to answer each other's questions.

@@

\section{Future Goals}
While much of my teaching experience has been with introductory
programming courses, I am also interested in teaching higher level
and theory based courses. Artificial intelligence and machine learning
are common courses which coincide strongly with my research interests.
Algorithms, discrete math, and graph theory are topics that I believe
present exciting challenges for keeping students engaged.
From previous work I have also acquired significant knowledge of
introductory and high performance C++ and would enjoy teaching classes
on those topics.
